\ttitle

### 2013-02-01 Chapters/Chapter2.tex
Fix typs
Detailed method of 'Segmenting Motion Capture Data into Distinct Behaviors'

\subsection{Formal definition}
Formally, temporal segmentation is dividing a time series $s$, which consists 
of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from 
$\mathbf{R}^d$. Individual \emph{segments} are referenced by $s(a,b)$, 
consisting of the consecutive samples 
$\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$. Let $s_1 = 
s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is 
$s_1s_2 = s(a,c)$. A segmentation $S$ of $s$ consists of a sequence of $k$ 
non-empty segments $s_1s_2 \dots s_k = s$.
This notation has close resemblance to the notation as used in 
\cite{himberg2001time}.

@@

@@

Given a set of data points, a lower dimensional hyperplane can by constructed 
to which the data points can be projected. This projection on a lower 
dimension introduces a error to the original position. When the error is 
fixed, less dimensions are needed for simple motions in which movements of 
body parts are highly correlated. For segments in which the data points are 
lesser correlated, e.g. because of transition state, a higher degree of 
dimensions of the hyperplane is needed to represent the data with equal error 
degree.

[SKIPPING]
By analyzing the derivative of the error rate, the algorithm is capable of 
selecting transition points and thus segmenting the data. When a data point 
$d_i$ is more than $k_\sigma=3$ times the standard deviation from the average 
over the previous data points $d_j, j<i$, then a cut is assigned to that frame.

@@

>>>>

\subsection{Formal definition}
Formally, temporal segmentation is dividing a time series $s$, which consists 
of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from 
$\mathbf{R}^d$. Individual \emph{segments} are referenced by $s(a,b)$, 
consisting of the consecutive samples 
$\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$. Let $s_1 = 
s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is 
$s_1s_2 = s(a,c)$. A segmentation $S$ of $s$ consists of a sequence of $k$ 
non-empty segments $s_1s_2 \dots s_k = s$.
This notation is adopted from \cite{himberg2001time}.

@@

\subsection{PCA Based Methods}

@@

Given a set of data points, a lower dimensional hyperplane can by constructed 
to which the data points can be projected. This projection on a lower 
dimension introduces a error to the original position. When the error is 
fixed, less dimensions are needed for simple motions in which movements of 
body parts are highly correlated. For segments in which the data points are 
lesser correlated, e.g. because of transition state, a higher degree of 
dimensions of the hyperplane is needed to represent the data with equal error 
degree. When the dimensionality is reduced from $d$ to $r \le d$, the ratio of 
error 
$E_r$ can be calculated as

\begin{equation}
	E_r = \frac{\sum_{j=1}^{r} \sigma_j^2}{\sum_{j=1}^{d} \sigma_j^2}
\end{equation}

where $\sigma_j$ are the singular values as a result from Singular Value 
Decomposition (SVD), which is closely related to PCA \cite{shlens2005tutorial}.

In \cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames 
to find transitions between action. First, for a number of $k$ frames (e.g. 
the 
equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error 
$E_r$ 
below some threshold $\tau$ is calculated. This will yield in an error $e_i$ 
for the first 
$i$ frames. When more frames are added to the window, the error will increase 
with a low constant when it is still in the same activity, due to noise in the 
activity. When a new activity 
starts, e.g. at frame $j$, the error at frame $j$ will start increasing 
faster. This can be expressed by the derivative of the error rate $d_i = e_i - 
e_{i-l}$, where $l$ is a constant to remove noise. From this derivative the 
mean and standard deviation can be calculated, for each point. When a 
derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations 
from the mean, a transition point is encountered. The previous frames are then 
cut from the sequence (as a segment) and the algorithm starts over.

*** [figure to illustrate derivative and standard deviation]

@@

*** [figure to illustrate mahalanobis distance measure and peaks]
